Metropolitan Statistical Area Evaluation

### 2023-04-20 src/demo_data_wrangling.Rmd
cleaning up code

### Population Growth
All states were considered with the exception of CA, HI, and AK. Initial analysis of Metropolitan Statistical Areas (MSAs) relied on Census data of annual population estimates, in which the population had to be greater than 500,000 individuals for the purpose of attracting investors. Population change from 2020 to 2021 was calculated by $$\frac{2021 estimate- 2020 estimate}{2020 estimate} * 100$$
MSAs with a population growth > 0.5%, were selected to gather further data on. The 0.5% growth was determined.

@@

Figure 1. Population growth in Metropolitan Statistical Areas (n=29) in 2020-2021. Highlighted areas have a population > 300,000 individuals and a population growth rate greater than 0.5%. The colors of the MSAs represent incremental growth rates, in which low(green) is 0.5-1% increase, medium(blue) is 1-2% increase, and high(pink) is 2-3.5% increase. Data: U.S. Census Bureau, Population Division.

@@

### Household Income
Household Median Income data was also obtained from Census data at the MSA level and amended to existing MSAs with a positive population increase. Median is a better indicator of the population than the mean income values due to the positive skew from the highest incomes.

@@

### Employment Rates
MSA monthly unemployment rates (seasonally-adjusted) were provided by The Bureau of Labor statistics in the civil labor force data at the MSA level. The average annual unemployment rates were calculated using the monthly rates from 2012 to 2020. Employment rates were calculated by dividing the number of employment / civilian labor force. The employment data for the month of January in years 2018-2021 was selected to calculate the change in employment rates over time. Year-over-Year employment growth was calculated by dividing the employment rate of the year by the previous year and subtracting one; for example, 2018 employment YoY growth = (2019 employment rate / 2018 employment rate) - 1.

@@

<https://www.bls.gov/web/metro.supp.toc.htm> <https://www.bls.gov/web/metro/metro_oty_change.htm> - significant change

2021 Median Rent from https://www.huduser.gov/portal/datasets/50per.html#2021

@@

### City Data Table
MSAs were split up by cities to collect utility, rent, and landlord data.

>>>>

### Population Growth

All states were considered with the exception of CA, HI, and AK. Initial analysis of Metropolitan Statistical Areas (MSAs) relied on Census data of annual population estimates, in which the population had to be greater than 500,000 individuals for the purpose of attracting investors. Population change from 2020 to 2021 was calculated by $$\frac{2021 estimate- 2020 estimate}{2020 estimate} * 100$$ MSAs with a population growth \> 0.5%, were selected to gather further data on. The 0.5% growth was determined.

@@

Figure 1. Population growth in Metropolitan Statistical Areas (n=29) in 2020-2021. Highlighted areas have a population \> 300,000 individuals and a population growth rate greater than 0.5%. The colors of the MSAs represent incremental growth rates, in which low(green) is 0.5-1% increase, medium(blue) is 1-2% increase, and high(pink) is 2-3.5% increase. Data: U.S. Census Bureau, Population Division.

@@

### Household Income

Household Median Income data was also obtained from Census data at the MSA level and amended to existing MSAs with a positive population increase. Median is a better indicator of the population than the mean income values due to the positive skew from the highest incomes.

@@

### Employment Rates

MSA monthly unemployment rates (seasonally-adjusted) were provided by The Bureau of Labor statistics in the civil labor force data at the MSA level. The average annual unemployment rates were calculated using the monthly rates from 2012 to 2020. Employment rates were calculated by dividing the number of employment / civilian labor force. The employment data for the month of January in years 2018-2021 was selected to calculate the change in employment rates over time. Year-over-Year employment growth was calculated by dividing the employment rate of the year by the previous year and subtracting one; for example, 2018 employment growth = (2019 employment rate / 2018 employment rate) - 1.

@@

2021 Median Rent from <https://www.huduser.gov/portal/datasets/50per.html#2021>

@@

### City Data Table

MSAs were split up by cities to collect utility, rent, and landlord data.
